Write-and-f-array: implementation and an application

### 2013-08-31 msc.tex
unified access to shared memory, rd and wr and cas used everywhere

\begin{algorithmic}[1]
	\State\var{$T$}{\arrayspec{P}$\int$} \Comment{T is an array of $P$ CAS objects of type \int}
\Function{inc$_p$}{}
	\State $x \gets T[p].\fn{rd}()$
	\State $T[p].\fn{wr}(x + 1)$
\EndFunction
\Function{read}{}
	\State $R \gets 0$
	\For{$i \gets 1 \mbox{to} P$}
		\State $R \gets R + T[i]$
	\EndFor
	\State\Return $R$
\EndFunction
\end{algorithmic}

@@

\begin{algorithmic}[1]
	\State\var{$S$}{$\textbf{pair}\left<\int v, \int p\right>$} \Comment{Version number and $p$ of the most recently published value}
	\State\var{$H$}{\arrayspec{N}$\textbf{pair}\left<\int v, \T T\right>$} \Comment{A circular buffer of last $N$ values (except maybe the most recent one)}
	\State\var{$L$}{\arrayspec{P}$\textbf{pair}\left<\int v, \T T\right>$} \Comment{Temporary storage for values being published}
	\Function{get-current}{}
		\State $s \gets S$ \Comment{Possible serialization point}
		\State \fn{help}()
		\State \Return $H[s.v \mod N]$
	\EndFunction
	\Function{help}{}
		\State $s \gets S$
		\State $l \gets L[s.p]$
		\State $h \gets H[s.v \mod N]$
		\If{$l.v = s.v \textbf{and} h.v < s.v$}
		\State $\text{CAS}(H[s.v \mod N], h, l)$ \label{hist-help-cas}
		\EndIf
	\EndFunction
	\Function{get}{v}
		\State $s \gets S$ \label{hist-get-notyet-sp} \Comment{Possible serialization point}
		\If{$s.v < v$}
		\State \Return \none \label{hist-get-notyet} \Comment{The requested version hasn't been published yet}
		\EndIf
		\State \fn{help}()
		\State $h \gets H[v \mod N]$
		\If{$h.v = v$}
			\State \Return $h.T$
		\Else
			\State \Return \none \label{hist-get-tooold}
		\EndIf
	\EndFunction
	\Function{publish$_p$}{v, T}
		\State $s \gets S$ \label{hist-pub-earlier-sp} \Comment{Possible serialization point}
		\If{$v \neq s.v + 1$}
			\State \Return \textbf{false} \label{hist-pub-earlier-exit}
		\EndIf
		\State \fn{help}()
		\State $L[p] \gets (v, T)$ \label{hist-pub-latest}
		\If{$\text{CAS}(S, s, (v, p)$) \textbf{failed}} \Comment{Possible serialization point}\label{hist-pub-later-sp}
			\State \Return \textbf{false}
		\EndIf
		\State \Return \textbf{true}
	\EndFunction
\end{algorithmic}

@@

\begin{Listing}[!h]
\begin{algorithmic}[1]
	\State\var{S}{\textbf{pair<int v, \T T>}}
	\Function{write-and-f$_0$}{T'}
		\State v, T $\gets$ S
		\State S $\gets$ (v + 1, T')
		\State \Return v + 1, v + 1, T
	\EndFunction
	\Function{get-last}{$i$}
		\State v, T $\gets$ S
		\State \Return v, v, T
	\EndFunction
	\Function{read}{}
		\State \Return S
	\EndFunction
\end{algorithmic}
\caption{The enriched write-and-f-array snapshot implementation for $N=1$}
\label{impl1}
\end{Listing}
\FloatBarrier

@@

\begin{algorithmic}[1]
	\algrestore{impl}
	\Function{help}{x}
		\State $l_c \gets C[side(x)].\fn{get-last}(child\_id(x))$ \label{get-last-recurse}
		\State $v_{current} \gets H.\fn{get-current}().v$
		\State binary search $\{v_{current}-(N+1),\ldots, v_{current}\}$ for first such $v$ that: \label{help-binsch}
		\Statex $H.\fn{get}(v) \neq \none \textbf{ and } H.\fn{get}(v).v[side(x)] \geq l_c.v$ %indent!
		\If{no such $v$ exists}
			\State \Return
		\EndIf
		\State $h_{old} \gets H.\fn{get}(v-1)$ \label{help-getprev}
		\State $h_{new} \gets H.\fn{get}(v)$
		\If{$h_{old} = \none \textbf{ or } h_{new} = \none$}
			\State \Return
		\EndIf
		\If{$side(x) = 0$}
		\State $T \gets f(l_c.T, h_{old}.T[1])$
		\Else
		\State $T \gets f(h_{new}.T[0], l_c.T)$
		\EndIf
		\State $l \gets L[x]$
		\If{$l.n < l_C.n$} \label{upd-last-alreadydone}
			\State $\text{CAS}(L[x], l, (l_C.n, v, T))$
		\EndIf
	\EndFunction
	\Function{get-last}{x}
		\State $\fn{help}(x)$
		\State \Return $L[x]$
	\EndFunction
\end{algorithmic}

>>>>

\begin{algorithmic}[1]
	\State\var{$T$}{\arrayspec{P}$\int$} \Comment{T is an array of $P$ CAS objects of type \int}
\Function{inc$_p$}{}
	\State $x \gets T[p].\fn{rd}()$
	\State $T[p].\fn{wr}(x + 1)$
\EndFunction
\Function{read}{}
	\State $R \gets 0$
	\For{$i \gets 1 \mbox{to} P$}
	\State $R \gets R + T[i].\fn{rd}$
	\EndFor
	\State\Return $R$
\EndFunction
\end{algorithmic}

@@

\begin{algorithmic}[1]
	\State\var{$S$}{$\textbf{pair}\left<\int v, \int p\right>$} \Comment{Version number and $p$ of the most recently published value}
	\State\var{$H$}{\arrayspec{N}$\textbf{pair}\left<\int v, \T T\right>$} \Comment{A circular buffer of last $N$ values (except maybe the most recent one)}
	\State\var{$L$}{\arrayspec{P}$\textbf{pair}\left<\int v, \T T\right>$} \Comment{Temporary storage for values being published}
	\Function{get-current}{}
		\State $s \gets S.\fn{rd}$ \Comment{Possible serialization point}
		\State \fn{help}()
		\State \Return $H[s.v \mod N].\fn{rd}$
	\EndFunction
	\Function{help}{}
		\State $s \gets S.\fn{rd}$
		\State $l \gets L[s.p].\fn{rd}$
		\State $h \gets H[s.v \mod N].\fn{rd}$
		\If{$l.v = s.v \textbf{and} h.v < s.v$}
		\State $H[s.v \mod N].\fn{CAS}(h, l)$ \label{hist-help-cas}
		\EndIf
	\EndFunction
	\Function{get}{v}
	\State $s \gets S.\fn{rd}$ \label{hist-get-notyet-sp} \Comment{Possible serialization point}
		\If{$s.v < v$}
		\State \Return \none \label{hist-get-notyet} \Comment{The requested version hasn't been published yet}
		\EndIf
		\State \fn{help}()
		\State $h \gets H[v \mod N].\fn{rd}$
		\If{$h.v = v$}
			\State \Return $h.T$
		\Else
			\State \Return \none \label{hist-get-tooold}
		\EndIf
	\EndFunction
	\Function{publish$_p$}{v, T}
		\State $s \gets S.\fn{rd}$ \label{hist-pub-earlier-sp} \Comment{Possible serialization point}
		\If{$v \neq s.v + 1$}
			\State \Return \textbf{false} \label{hist-pub-earlier-exit}
		\EndIf
		\State \fn{help}()
		\State $L[p].\fn{wr}((v, T))$ \label{hist-pub-latest}
		\If{$S.\fn{CAS}(s, (v, p)$) \textbf{failed}} \Comment{Possible serialization point}\label{hist-pub-later-sp}
			\State \Return \textbf{false}
		\EndIf
		\State \Return \textbf{true}
	\EndFunction
\end{algorithmic}

@@

\begin{Listing}[!h]
\begin{algorithmic}[1]
	\State\var{S}{\textbf{pair<int v, \T T>}} %TODO: local var name T is misleading. rename?
	\Function{write-and-f$_0$}{T'}
		\State $v, T \gets S.\fn{rd}$
		\State $S.\fn{wr}((v + 1, T'))$
		\State \Return $v + 1, v + 1, T$
	\EndFunction
	\Function{get-last}{$i$}
		\State $v, T \gets S.\fn{rd}$
		\State \Return $v, v, T$
	\EndFunction
	\Function{read}{}
		\State \Return $S.\fn{rd}$
	\EndFunction
\end{algorithmic}
\caption{The enriched write-and-f-array snapshot implementation for $N=1$}
\label{impl1}
\end{Listing}
\FloatBarrier

@@

\begin{algorithmic}[1]
	\algrestore{impl}
	\Function{help}{x}
		\State $l_c \gets C[side(x)].\fn{get-last}(child\_id(x))$ \label{get-last-recurse}
		\State $v_{current} \gets H.\fn{get-current}().v$
		\State binary search $\{v_{current}-(N+1),\ldots, v_{current}\}$ for first such $v$ that: \label{help-binsch}
		\Statex $H.\fn{get}(v) \neq \none \textbf{ and } H.\fn{get}(v).v[side(x)] \geq l_c.v$ %indent!
		\If{no such $v$ exists}
			\State \Return
		\EndIf
		\State $h_{old} \gets H.\fn{get}(v-1)$ \label{help-getprev}
		\State $h_{new} \gets H.\fn{get}(v)$
		\If{$h_{old} = \none \textbf{ or } h_{new} = \none$}
			\State \Return
		\EndIf
		\If{$side(x) = 0$}
		\State $T \gets f(l_c.T, h_{old}.T[1])$
		\Else
		\State $T \gets f(h_{new}.T[0], l_c.T)$
		\EndIf
		\State $l \gets L[x].\fn{rd}$
		\If{$l.n < l_C.n$} \label{upd-last-alreadydone}
			\State $L[x].\text{CAS}(l, (l_C.n, v, T))$
		\EndIf
	\EndFunction
	\Function{get-last}{x}
		\State $\fn{help}(x)$
		\State \Return $L[x].\fn{rd}$
	\EndFunction
\end{algorithmic}
